MR-PFU: A suite of R packages for energy conversion chain analysis

### 2023-07-24 paper.Rmd
Add DeSterck reference

We created a suite of `R` packages 
(some of which are generally useful and available through 
[CRAN][link-cran],
others of which are focused on SEA and available on GitHub)
to assist SEA practitioners as they analyze 
energy movement through society. 
These new packages enable analysis of every country in the world
across timespans of decades or longer.
The new packages were used to create the "PFU Database," 
a new resource for the SEA community. 
The packages can be used for specific SEA analyses, too.

@@

Historically, societal exergy analysts have
analyzed the ECCs of individual countries
using a large number of linked spreadsheets,
often starting with primary- and final-stage data 
from the [IEA][link-ieaweeb]'s
world energy balances
[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
Data were in varying and inconsistent formats.
The authors of this paper and others in the field
wanted to expand SEA to cover all countries, but
the spreadsheet approach to SEA was
deemed not scalable.
We created a suite of `R` packages to create a database^[Strictly speaking, we create data frames of matrices, 
not a structured database.]
of PFU data (the PFU Database)
for all countries in the world from 1960--2020.
The packages enable, for the first time, scalable SEA.
This paper describes the design of those packages and
demonstrates briefly their use.

>>>>

We created a suite of `R` packages 
(some of which are generally useful and available through 
[CRAN][link-cran],
others of which are focused on SEA and available on GitHub)
to assist SEA practitioners as they analyze 
energy movement through society. 
The new packages enable analysis of any country in the world
across timespans of decades or longer.
In short, the packages enable, for the first time, scalable SEA.
We used the new packages to create the "PFU Database," 
a new resource for the SEA community^[Strictly speaking, 
we create data frames of matrices, 
not an SQL or similar database.].

This paper describes the design of the new packages and
demonstrates briefly their use.

@@

Historically, societal exergy analysts have
analyzed the ECCs of individual countries
using a large number of linked spreadsheets,
often starting with primary- and final-stage data 
from the [IEA][link-ieaweeb]'s
world energy balances
[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
Data were stored in varying and inconsistent formats.
A different attempt to create an SEA database
estimated technical energy efficiencies 
of end-use machines by economic status in a country,
thereby precluding the use of the database for energy-economy studies [@De-Stercke:2014].

The authors of this paper and others in the field
wanted to expand SEA to cover all countries, but
the spreadsheet approach to SEA was
deemed not scalable.
A new approach to SEA was needed, 
one that scaled across all countries 
for many years without relying on economic data
to estimate technical efficiencies.
